An Environment for Sustainable Research Software in Germany and Beyond: Current State, Open Challenges, and Call for Action

### 2020-04-27 001/manuscript.tex
Update PDF

%%% Journal toggle; only specific options recognised.
%%% (Only "gigascience" and "general" are implemented now. Support for other journals is planned.)
\reviewfrom{23 January 2020}
\reviewto{09 February 2020}
\acceptedon{03 April 2020}

@@

We consider Software Carpentry and similar efforts like the creation of the Data Science Academy HIDA~\cite{HIDA} in the Helmholtz Association of German Research Centers helpful solutions to exchange and distribute knowledge. Local chapters of RSE groups and (inter-)national conferences will further foster networking and community building. We strongly recommend the creation of a national Software Sustainability Institute (involving funded positions to establish web platforms and training material) similar to the \cite{UKSoftwareSustainabilityInstitute}, which serves as a national contact for all aspects related to research software. The UK SSI also publishes best practice guidelines~\cite{SoftwareSystemsDevelopmentLifeCycle_SDLC} for research software engineering.

@@

\subsection{Archiving}
Software preservation aims to extend the lifetime of software that is no longer actively maintained. There are different approaches, which vary in the effort required and the likelihood of success. Software archiving is one important aspect of software preservation: the process of storing a copy of a software so that it may be referred to in the future.
The publication of a certain software version for reference in research articles requires simple ways to archive research software on a long-term basis. Furthermore, its integration with collaborative software development environments such as GitLab or GitHub and with publication repositories is needed to facilitate archiving of referenced software versions based on sustainable frameworks (e.g. Invenio~\cite{Invenio}) for GitHub to Zenodo integration).

>>>>

%%% Journal toggle; only specific options recognised.
%%% (Only "gigascience" and "general" are implemented now. Support for other journals is planned.)
\reviewfrom{23 January 2020}
\reviewto{09 February 2020}
\acceptedon{03 April 2020}
\publicationurl{tba}
\publicationpidtype{arXiv}
\publicationpid{submit/3149954}

@@

We consider Software Carpentry and similar efforts like the creation of the Data Science Academy HIDA~\cite{HIDA} in the Helmholtz Association of German Research Centers helpful solutions to exchange and distribute knowledge. Local chapters of RSE groups and (inter-)national conferences will further foster networking and community building. We strongly recommend the creation of a national Software Sustainability Institute (involving funded positions to establish web platforms and training material) similar to the existing institute in the UK~\cite{UKSoftwareSustainabilityInstitute}, which serves as a national contact for all aspects related to research software. The UK SSI also publishes best practice guidelines~\cite{SoftwareSystemsDevelopmentLifeCycle_SDLC} for research software engineering.

@@

\subsection{Archiving}
Software preservation aims to extend the lifetime of software that is no longer actively maintained. There are different approaches, which vary in the effort required and the likelihood of success. Software archiving is one important aspect of software preservation: the process of storing a copy of a software so that it may be referred to in the future.
The publication of a certain software version for reference in research articles requires simple ways to archive research software on a long-term basis. Furthermore, its integration with collaborative software development environments such as GitLab or GitHub and with publication repositories is needed to facilitate archiving of referenced software versions based on sustainable frameworks (e.g. Invenio~\cite{Invenio} for GitHub to Zenodo integration).
